Evolution, functional diversity, regime shifts, hysteresis, and nonlinearity; ecosystem responses to environmental change.

### 2021-07-05 manuscript/article.Rmd
figures and results to discuss

```{r include = FALSE}
knitr::opts_chunk$set(include = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = TRUE)
```

```{r}
library(tidyverse)
library(microxanox)
library(patchwork)
library(here)
source(here("simulations/r functions/various_useful_functions.r"))
```

@@

This report concerns how biodiversity influences the strength and nature of ecosystem responses to environmental change (figure 1; panel(a) three ovals, environment oval with arrow to ecosystem property oval, with biodiversity oval pointing the arrow between environment and ecosystem property; panel (b) shows linear response with different gradients, large being with low diversity; panel (c) shows a linear and a non-linear response starting and ending at the same place; panel (d) shows a hysteretic response, with greater or less range of hysteresis). Strong responses of ecosystems to environmental change imply high *sensitivity* and lack of resistance to environmental change, and higher temporal variability in the face of varying environmental conditions. Greater *non-linearity* in ecosystem responses increases the potential for surprises, such as when an ecosystem that was resistant to an environmental change becomes unexpectedly sensitive to further change. The presence and extent of *hysteresis* in ecosystem responses to environmental driver gradients concerns how readily are ecosystem responses reversed when environmental change is reversed.

@@

-   In general, greater functional response diversity may decrease the likelihood of a catastrophic ecosystem response (Figure 2a in @dakos2019b). Here, we interpret this prediction to be that greater functional response diversity decreases the non-linearity of ecosystem response to environmental driver change, including the extreme non-linearity that can occur when an ecosystem "flips" from one state to another one.

@@

These explanations/predictions were based on the verbal/graphical argumentation @dakos2019b. There were no mathematical models of ecosystem responses to environmental change behind the explanations/predictions. @ceulemans2021, in their article *Functional diversity alters the effects of a pulse perturbation on the dynamics of tritrophic food webs*, did employ a such a mathematical model [to be developed].

@@

* To provide an answer to the question of how trait variation affects the non-linearity, resilience and hysteresis of ecosystem response to an environmental change.

@@

* Mathematical modelling of a microbrial ecosystem. Trait diversity manipulation (can be considered selection among strains, or selection among species). (Do not include mutation or recomination).

* Variation in amount of among strain variation, no variation in number of strains, no variation in strength or shape of the of any tradeoff.

* Measurement of the effect on how the steady state(s) of the system varies with in different environmental conditions, and on the dynamical response of system to gradual environmental change.

@@

```{r ecosys_network, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("your caption"), echo=FALSE}
knitr::include_graphics(here("manuscript/ecosystem_network.png"))
```

@@

@@

In this section we describe how we created variability among the multiple strains within functional groups. Introducing variation among multiple strains within a functional group required a number of decisions:

* **The number of strains.** The symbol for this is $n$ subscripted with the functional group code (CB = cyanobacteria, SP = sulphate reducing bacteria, and PB = phototrophic sulphur bacteria). E.g. $n_{CB}$ is the number of strains of cyanobacteria. [add why we only ever had 9].
* **Which traits to introduce variation in.** We chose the maximum growth rate and the tolerance of growth to an inhibiting substrate. E.g. maximum growth rate of the cyanobacteria $G_{max,CB}$, and the tolerance of cyanobacterial growth to reduced sulphur $h_{SR,CB}$.
* **The amount of among strain variation.** The symbol for this is $Div$ (from *Div*ersity) subscripted by the trait/parameter and the functional group. E.g. * $Div_{G_{max,CB}}$  is the amount of among strain variation in the maximum growth rate $G_{max}$ of the cyanobacteria $CB$. 
* **The correlation between variation in different traits.** This was always negative, so there was a tradeoff. Our implementation of the variation and tradeoff contain the assumption of a linear tradeoff of log-log transformed among-strain variation.
* **The relative amount of among strain variation in different functional groups.** E.g. were strains of cyanobacteria very variable compared to strains of the other functional groups. We performed several numerical experiments, varying which functional groups had more or less among strain variation.

* $n_{CB}$ - number of strains of cyanobacteria.
* $n_{SB}$ - number of strains of sulphate reducing bacteria.
* $n_{PB}$ - number of strains of phototrophic sulphur bacteria.
* $Div_{G_{max},CB}$ - amount of among strain variation in the maximum growth rate $G_{max}$ of the cyanobacteria $CB$. 
* $Div_{h,CB}$ - amount of among strain variation in the maximum growth rate $G_{max}$ of the cyanobacteria $CB$. 
* $Div_{G_{max},SB}$ - amount of among strain variation in the maximum growth rate $G_{max}$ of the sulphate reducing bacteria $SB$. 
* $Div_{h,SB}$ - amount of among strain variation in the maximum growth rate $G_{max}$ of the sulphate reducing bacteria $SB$. 
* $Div_{G_{max},PB}$ - amount of among strain variation in the maximum growth rate $G_{max}$ of the phototrophic sulphur bacteria $PB$. 
* $Div_{h,PB}$ - amount of among strain variation in the maximum growth rate $G_{max}$ of the phototrophic sulphur bacteria $PB$.

We introduced among-strain variability in maximum growth rate (parameter $G_{max}$) and in the tolerance to the relevant inhibiting chemical substrate (parameter $h$). Among-strain variability was created such that there was a tradeoff between maximum growth rate and tolerance (i.e. high tolerance coinciding with low maximum growth rate) (e.g. Figure \@ref(fig:tradeoff). Our implementation of the variation and tradeoff contain the assumption of a linear tradeoff of log-log transformed among-strain variation.

```{r tradeoff, include = TRUE, echo = FALSE, out.width="0.5\\linewidth", fig.align="center", fig.cap=c("An example of among-strain variation in growth rate and tolerance, and the tradeoff in the variation.")}
ggplot() +
  geom_point(aes(x = G_max_i, y = h_i)) +
  xlab("Maximum growth rate") +
  ylab("Tolerance") +
  theme_bw(base_size = 20)
```

For further understanding, please consider a reference maximum growth rate parameter $G_{max}$ and reference tolerance parameter $h$. Furthermore, consider the meta-parameters $Div_{G_{max}}$ and $Div_{h}$, which determine the amount of variation among $n$-strains in the growth rate and tolerance parameters. To create a tradeoff, the two meta-parameters were always of different sign.

The among strain variation was calculated such that the growth rate of strain $i = 1$ was a factor $1/(2^{Div_{G_{max}}})$ of the reference growth rate, and the growth rate of the $i = n$ strain was a factor $2^{Div_{G_{max}}}$ of the reference growth rate. The growth rate of the $i = \{2,...n-1\}$ other strains was the reference growth rate multiplied by a factor that was equally distributed between the factor of the $i = 1$ and $i = n$ strain.

Hence, for example, with the reference growth rate $G_{max} =$ `r G_max`, among strain variation of $Div_{G_{max}} =$ `r Div_G_max`, and nine strains ($n =$ `r n`), then the growth rates of the nine strains are `r round(G_max_i,3)` (Figure \@ref(fig:tradeoff).

(Put another way, the growth rate parameter of the i-th of the $n$ strains in a functional group was calculated as $G_{max, i} = D_i * G_{max}$ where D_i is the i-th element of the set $D = \{2^{f(x)Div_{G_{max}}} | x \in \mathbb{N}, x \leq n\}$, where $f(x) = (x-(n+1)/2) / (( n-1)/2)$.)

@@

```{r ss_novar, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("No diversity"), echo=FALSE}
CBSBPB_data <- readRDS(here("simulations/expt2/data/ss_res_grad_CBSBPB_var.RDS"))
result_index1 <- 1
p1  <- plot_ss_result1(CBSBPB_data,
                result_index = result_index1,
                filename_prefix = NULL,
                save_image_file = FALSE)
p1

@@

```{r ss_var, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Diversity in all functional groups"), echo=FALSE}
result_index2 <- 20
p2  <- plot_ss_result1(CBSBPB_data,
                result_index = result_index2,
                filename_prefix = NULL,
                save_image_file = FALSE)
p2

@@

```{r ss_var2, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Diversity in none (solid) and all functional groups."), echo=FALSE}
p_overlay <- plot_ss_result2(CBSBPB_data[result_index1,]$ss_res[[1]],
                             CBSBPB_data[result_index2,]$ss_res[[1]],
                             xlims = c(-7, -1))
p_overlay
```

@@

```{r CB_var1, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Diversity in CB functional group."), echo=FALSE}
CB_data <- readRDS(here("simulations/expt1/data/ss_res_grad_CB_var.RDS"))
p3  <- plot_ss_result1(CB_data,
                result_index = result_index2,
                filename_prefix = NULL,
                save_image_file = FALSE)
p3
```

```{r CB_var2, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Diversity in none (solid) and CB functional groups."), echo=FALSE}
p_overlay <- plot_ss_result2(CB_data[result_index1,]$ss_res[[1]],
                             CB_data[result_index2,]$ss_res[[1]],
                             xlims = c(-7, -1))
p_overlay
```

## Figure: SB Stable states with variation and tradeoff

```{r SB_var1, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Diversity in SB functional group."), echo=FALSE}
SB_data <- readRDS(here("simulations/expt1/data/ss_res_grad_SB_var.RDS"))
p4  <- plot_ss_result1(SB_data,
                result_index = result_index2,
                filename_prefix = NULL,
                save_image_file = FALSE)
p4

@@

```{r SB_var2, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Diversity in none (solid) and SB functional group."), echo=FALSE}
p_overlay <- plot_ss_result2(CB_data[result_index1,]$ss_res[[1]],
                             SB_data[result_index2,]$ss_res[[1]],
                             xlims = c(-7, -1))

## Figure: PB Stable states with variation and tradeoff

```{r PB_var1, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Diversity in PB functional group."), echo=FALSE}
PB_data <- readRDS(here("simulations/expt1/data/ss_res_grad_PB_var.RDS"))
p5  <- plot_ss_result1(PB_data,
                result_index = result_index2,
                filename_prefix = NULL,
                save_image_file = FALSE)
p5

```{r PB_var2, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Diversity in none (solid) and PB functional group."), echo=FALSE}
p_overlay <- plot_ss_result2(CB_data[result_index1,]$ss_res[[1]],
                             PB_data[result_index2,]$ss_res[[1]],
                             xlims = c(-7, -1))

```{r SBPB_var1, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Diversity in SB and PB functional groups."), echo=FALSE}
SBPB_data <- readRDS(here("simulations/expt1/data/ss_res_grad_SBPB_var.RDS"))
p6  <- plot_ss_result1(SBPB_data,
                result_index = result_index2,
                filename_prefix = NULL,
                save_image_file = FALSE)
p6

```{r SBPB_var2, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Diversity in none (solid) and SB and PB functional groups."), echo=FALSE}
p_overlay <- plot_ss_result2(CB_data[result_index1,]$ss_res[[1]],
                             SBPB_data[result_index2,]$ss_res[[1]],
                             xlims = c(-7, -1))

@@

```{r echo=FALSE, eval = FALSE}
#ss_sim_data <- readRDS(here("simulations/expt1/data/ss_res_3.RDS"))

CBSBPB_stab_data <- CBSBPB_data %>%
  group_by(CB_var_gmax_s, CB_var_h_s,
           SB_var_gmax_s, SB_var_h_s,
           PB_var_gmax_s, PB_var_h_s) %>%
  do(stability_measures = get_stability_measures(.$ss_res[[1]]))
results1 <- unnest(CBSBPB_stab_data, cols = c(stability_measures)) %>%
  mutate(var_treat = "CB-SB-PB",
         var_gmax = CB_var_gmax_s)

CB_stab_data <- CB_data %>%
  group_by(CB_var_gmax_s, CB_var_h_s,
           SB_var_gmax_s, SB_var_h_s,
           PB_var_gmax_s, PB_var_h_s) %>%
  do(stability_measures = get_stability_measures(.$ss_res[[1]]))
results2 <- unnest(CB_stab_data, cols = c(stability_measures))  %>%
  mutate(var_treat = "CB",
         var_gmax = CB_var_gmax_s)

SB_stab_data <- SB_data %>%
  group_by(CB_var_gmax_s, CB_var_h_s,
           SB_var_gmax_s, SB_var_h_s,
           PB_var_gmax_s, PB_var_h_s) %>%
  do(stability_measures = get_stability_measures(.$ss_res[[1]]))

# for(i in 1:20) {
#   print(i)
#   junk <- get_stability_measures(SB_data$ss_res[[i]])
# }
# ss_object <- SB_data$ss_res[[11]]
# 
# temp2 <- filter(temp, Species == "PB_1")
# get_hysteresis_total(log10(temp2$up+1), log10(temp2$down+1))
# down <- log10(temp2$down+1)
# up <- log10(temp2$up+1)
#

results3 <- unnest(SB_stab_data, cols = c(stability_measures))  %>%
  mutate(var_treat = "SB",
         var_gmax = SB_var_gmax_s)

PB_stab_data <- PB_data %>%
  group_by(CB_var_gmax_s, CB_var_h_s,
           SB_var_gmax_s, SB_var_h_s,
           PB_var_gmax_s, PB_var_h_s) %>%
  do(stability_measures = get_stability_measures(.$ss_res[[1]]))
results4 <- unnest(PB_stab_data, cols = c(stability_measures))  %>%
  mutate(var_treat = "PB",
         var_gmax = PB_var_gmax_s)

SBPB_stab_data <- SBPB_data %>%
  group_by(CB_var_gmax_s, CB_var_h_s,
           SB_var_gmax_s, SB_var_h_s,
           PB_var_gmax_s, PB_var_h_s) %>%
  do(stability_measures = get_stability_measures(.$ss_res[[1]]))
results5 <- unnest(SBPB_stab_data, cols = c(stability_measures))  %>%
  mutate(var_treat = "SB-PB",
         var_gmax = SB_var_gmax_s)

all_stab_results <- results1 %>%
  bind_rows(results2) %>%
  bind_rows(results3) %>%
  bind_rows(results4) %>%
  bind_rows(results5)

all_stab_results<- all_stab_results %>%
  mutate(var_treat = forcats::fct_relevel(var_treat, levels = c("CB", "SB", "PB", "SB-PB", "CB-SB-PB")))

saveRDS(all_stab_results, here("simulations/sim data/stab_measures.RDS"))

@@

```{r ss_var3, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("your caption"), echo=FALSE}

all_stab_results <- readRDS(here("simulations/expt1/data/stab_measures.RDS"))

all_stab_results %>%
  filter(Species == "O") %>%
  ggplot(aes(x = var_gmax, y = hyst_range, col=var_treat)) +
  geom_point() +
  xlab("Amount of trait variation\n[see text for units]") +
  ylab("Extent of bistability region\n[log10 oxygen diffusivity]") +
  labs(col = "Community\ncomposition")
```

```{r ss_var4, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("your caption"), echo=FALSE}
all_stab_results %>%
  #filter(var_treat == "CB") %>%
  filter(Species == "O") %>%
  ggplot(aes(x = var_gmax,
             ymin = hyst_min,
             ymax = hyst_max,
             fill=var_treat)) +
  geom_ribbon(alpha = 0.5) +
  facet_wrap( ~ var_treat) +
  xlab("Amount of trait variation\n[see text for units]") +
  ylab("Oxygen diffusivity\n[log10 uM per hour]") +
  labs(fill = "Community\ncomposition")

@@

## Figure: Temporal switches

Maybe... to be added. Need to go both ways (oxic to anoxic, and anoxic to oxic).

@@

>>>>

```{r include = FALSE}
knitr::opts_chunk$set(include = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE)
```

```{r}
library(tidyverse)
library(microxanox)
library(patchwork)
library(here)
source(here("experiments/r functions/various_useful_functions.r"))
```

@@

This report concerns how biodiversity influences the strength and nature of ecosystem responses to environmental change (figure 1; panel(a) three ovals, environment oval with arrow to ecosystem property oval, with biodiversity oval pointing the arrow between environment and ecosystem property; panel (b) shows linear response with different gradients, large being with low diversity; panel (c) shows a linear and a non-linear response starting and ending at the same place; panel (d) shows a hysteretic response, with greater or less range of hysteresis). Strong responses of ecosystems to environmental change imply high sensitivity and lack of resistance to environmental change, and higher temporal variability in the face of varying environmental conditions. Greater non-linearity in ecosystem responses increases the potential for surprises, such as when an ecosystem that was resistant to an environmental change becomes unexpectedly sensitive to further change. The presence and extent of hysteresis in ecosystem responses to environmental driver gradients concerns how readily are ecosystem responses reversed when environmental change is reversed.

@@

-   Greater functional response diversity may decrease the likelihood of a catastrophic ecosystem response (Figure 2a in @dakos2019b). Here, we interpret this prediction to be that greater functional response diversity decreases the non-linearity of ecosystem response to environmental driver change, including the extreme non-linearity that can occur when an ecosystem "flips" from one state to another one.

@@

These explanations/predictions were based on the verbal and graphical argumentation @dakos2019b. There were no mathematical models of ecosystem responses to environmental change behind the explanations/predictions. @ceulemans2021, in their article *Functional diversity alters the effects of a pulse perturbation on the dynamics of tritrophic food webs*, did employ a such a mathematical model [text to be developed].

@@

* To provide an answer to the question of how trait variation affects the resilience and hysteresis of ecosystem response to an environmental change.

@@

* Mathematical modelling of a microbial ecosystem.

* Trait diversity manipulation (can be considered selection among strains, or selection among species). (Do not include mutation or recombination).

* Variation in amount of among strain variation in two traits, no variation in number of strains, no variation in strength or shape of the of the tradeoff.

* Illustration of the dynamical response to gradual environmental change.

* Measurement of the effect trait diversity on how the steady state(s) of the system varies with environmental conditions.

@@

```{r ecosys_network, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("your caption"), echo=FALSE}
knitr::include_graphics(here("manuscript/figures/ecosystem_network.png"))
```

@@

```{r diversity_examples, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("your caption"), echo=FALSE}
knitr::include_graphics(here("manuscript/figures/diversity_example.pdf"))
```

In this section we describe how we created variability among the multiple strains within functional groups. In brief, in each functional group we created multiple distinct strains that differed in their maximum growth rate and tolerance to an inhibiting substrate, with a trade-off between the two (e.g. higher tolerance assocated with lower maximum growth rate). An example of the resulting variation is displayed in Figure \ref(fig:diversity_example)).

A more detailed account of the creation of strains follows now. Introducing variation among multiple strains within a functional group required a number of decisions:

* **The number of strains.** The symbol for this is $n$ subscripted with the functional group code (CB = cyanobacteria, SP = sulphate reducing bacteria, and PB = phototrophic sulphur bacteria). E.g. $n_{CB}$ is the number of strains of cyanobacteria. [add why we only ever had 9].
* **Which traits to introduce variation in.** We chose the maximum growth rate and the tolerance of growth to an inhibiting substrate. E.g. maximum growth rate of the cyanobacteria $G_{max,CB}$, and the tolerance of cyanobacterial growth to reduced sulphur $h_{SR,CB}$.
* **The amount of among strain variation.** The symbol for this is $Div$ (from *Div*ersity) subscripted by the trait/parameter and the functional group. E.g. * $Div_{G_{max,CB}}$  is the amount of among strain variation in the maximum growth rate $G_{max}$ of the cyanobacteria $CB$. 
* **The correlation between variation in different traits.** This was always negative, so there was a tradeoff. Our implementation of the variation and tradeoff contain the assumption of a linear tradeoff of log-log transformed among-strain variation.
* **The relative amount of among strain variation in different functional groups.** E.g. were strains of cyanobacteria very variable compared to strains of the other functional groups. We performed several numerical experiments, varying which functional groups had more or less among strain variation.
* **The distribution of trait variation.** We chose to have among strain variation distributed evenly on a log scale.

The parameters therefore necessary were:

* $n_{CB}$ - number of strains of cyanobacteria.
* $n_{SB}$ - number of strains of sulphate reducing bacteria.
* $n_{PB}$ - number of strains of phototrophic sulphur bacteria.
* $Div_{G_{max,CB}}$ - amount of among strain variation in the maximum growth rate $G_{max}$ of the cyanobacteria $CB$. 
* $Div_{h_{SR,CB}}$ - amount of among strain variation in the tolerance $h$ of cyanobacteria $CB$ to reduced sulpur compounds $SR$. 
* $Div_{G_{max,SB}}$ - amount of among strain variation in the maximum growth rate $G_{max}$ of the sulphate reducing bacteria $SB$. 
* $Div_{h_{O,SB}}$ - amount of among strain variation in the tolerance $h$ of sulphate reducing bacteria $S$ to oxygen $O$. 
* $Div_{G_{max,PB}}$ - amount of among strain variation in the maximum growth rate $G_{max}$ of the phototrophic sulphur bacteria $PB$. 
* $Div_{h_{O,PB}}$ - amount of among strain variation in the tolerance $h$ of phototrophic sulphur bacteria $PB$ to oxygen $O$.

For example, cyanobacteria, and a reference maximum growth rate parameter $G_{max,CB}$ and reference tolerance parameter $h_{SR,CB}$. Recall that amount of among strain variation is controlled by the two meta-parameters $Div_{G_{max,CB}}$ and $Div_{h_{SR,CB}}$. To create a tradeoff, the two meta-parameters were always of different sign. The among strain variation was calculated such that the growth rate of strain $i = 1$ was a factor $1/(2^{Div_{G_{max,CB}}})$ of the reference growth rate, and the growth rate of the $i = n$ strain was a factor $2^{Div_{G_{max,CB}}}$ of the reference growth rate. The growth rate of the $i = \{2,...n-1\}$ other strains was the reference growth rate multiplied by a factor that was equally distributed between the factor of the $i = 1$ and $i = n$ strain. This implementation of the variation and trade-off contains the assumption of a linear trade-off of log-log transformed among-strain variation.

@@

```{r tradeoff, eval = FALSE, include = TRUE, echo = FALSE, out.width="0.5\\linewidth", fig.align="center", fig.cap=c("An example of among-strain variation in growth rate and tolerance, and the tradeoff in the variation.")}
ggplot() +
  geom_point(aes(x = G_max_i, y = h_i)) +
  xlab("Maximum growth rate") +
  ylab("Tolerance") +
  theme_bw(base_size = 20)
```

Hence, for example, with the reference growth rate $G_{max,CB} =$ `r G_max`, among strain variation of $Div_{G_{max,CB}} =$ `r Div_G_max`, and nine strains ($n =$ `r n`), then the growth rates of the nine strains are `r round(G_max_i,3)` (Figure to be added).

(Put another way, the growth rate parameter of the i-th of the $n$ strains in a functional group was calculated as $G_{max, i} = D_i * G_{max}$ where $D_i$ is the i-th element of the set $D = \{2^{f(x)Div_{G_{max}}} | x \in \mathbb{N}, x \leq n\}$, where $f(x) = (x-(n+1)/2) / (( n-1)/2)$.)

@@

```{r ss_novar, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem with no trait diversity, i.e. identical trait values for the strains within a functional group."), echo=FALSE}
CBSBPB_data <- readRDS(here("experiments/experiment 1/data/ss_data.RDS"))
result_index1 <- 1
p1  <- plot_ss_result1(CBSBPB_data,
                result_index = result_index1,
                filename_prefix = NULL,
                save_image_file = FALSE)
p1

@@

```{r ss_var, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem with diversity in all functional groups"), echo=FALSE}
result_index2 <- 20
p2  <- plot_ss_result1(CBSBPB_data,
                result_index = result_index2,
                filename_prefix = NULL,
                save_image_file = FALSE)
p2

@@

```{r ss_var2, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem without any diversity (solid line) and with diversity in all three functional groups."), echo=FALSE}
p_overlay <- plot_ss_result2(CBSBPB_data[result_index1,]$ss_res[[1]],
                             CBSBPB_data[result_index2,]$ss_res[[1]],
                             xlims = c(-7, -1))
p_overlay
```

@@

```{r CB_var1, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem with diversity in only the cyanobacteria functional group."), echo=FALSE}
CB_data <- readRDS(here("experiments/experiment 2/data/ss_data.RDS"))
p3  <- plot_ss_result1(CB_data,
                result_index = result_index2,
                filename_prefix = NULL,
                save_image_file = FALSE)
p3
```

```{r CB_var2, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem without any diversity (solid line) and with diversity in only the cyanobacteria functional group."), echo=FALSE}
p_overlay <- plot_ss_result2(CB_data[result_index1,]$ss_res[[1]],
                             CB_data[result_index2,]$ss_res[[1]],
                             xlims = c(-7, -1))
p_overlay
```

```{r SBPB_var1, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem with diversity in only the two sulphur bacteria functional groups."), echo=FALSE}
SBPB_data <- readRDS(here("experiments/experiment 3/data/ss_data.RDS"))
p6  <- plot_ss_result1(SBPB_data,
                result_index = result_index2,
                filename_prefix = NULL,
                save_image_file = FALSE)
p6

@@

```{r SBPB_var2, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Steady states of the ecosystem without any diversity (solid line) and with diversity in only the sulphur bacteria functional group."), echo=FALSE}
p_overlay <- plot_ss_result2(CB_data[result_index1,]$ss_res[[1]],
                             SBPB_data[result_index2,]$ss_res[[1]],
                             xlims = c(-7, -1))

@@

```{r stab1, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Effect of increasing amounts of trait diversity on the extent of the environmental gradient for which bistability (i.e. alternative stable states) is present, when different combinations of the three functional groups contain any diversity."), echo=FALSE}

knitr::include_graphics(here("manuscript/figures/extent_of_bistab1.pdf"))

@@

```{r stab2, out.width="1\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Effect of increasing amounts of trait diversity on the upper and lower bound of the environmental gradient for which bistability (i.e. alternative stable states) is present, when different combinations of the three functional groups contain any diversity."), echo=FALSE}

knitr::include_graphics(here("manuscript/figures/extent_of_bistab2.pdf"))

@@

@@

We observed that greater trait diversity among the organisms dominating the prevailing ecosystem state can lead to greater resilience of that state. We also observed that greater trait diversity among the organisms that occur in an different ecosystem state can reduce resilience.

The effect of diversity occurs because with greater trait diversity there is a organism with greater tolerance to the deteriorating (from the perspective of the dominant organisms). This greater tolerance means that a greater amount of environmental deterioration is required to cause the ecosystem to flip into the alternative state.

In the case study, with three functional groups, the presence or absence of diversity in the different functional groups influences resilience... with only diversity in the CB...
